### 2012-04-21 ReplyToReviewers.md
corrections made based on reviews from Interface

Replies in _italics_

Referee: 1
==========

### Comments to the Author
I think that this is one of the most important papers yet written in what is emerging as an area of great interest.  My only concern is that it is written in a highly technical way that may not be accessible to the broad audience that should read it.   The idea of using a model-based approach and approaching the problem as one of model choice is very clever, but I had to read the section three times before I knew what that meant.   I would move some of that material up front to make the main points a little less drily than they are currently made. I think also some technical terms should be embellished with discussion and explanation up front.

_We thank the reviewer for this assessment, and we have done our best to address this important issue_

Tell the reader, for example, what a saddle-node bifurcation is.

_A definition now follows the first mention_

Explain how the coefficient of the noise term in the stochastic differential equation is chosen.

_Certainly this needed more explanation for any audience.  The justification now follows where this appears_

This part is written for experts, whereas for Interfaces it needs to be more pedagogical.

Why does the bifurcation parameter have a subscript t?

_to indicate it is changing in time, this has been stated when it first appears_

### Small points:  
- Line 25:  lead should be led
- Line 298:  Data are
- Line 309:  its

_Indeed, thanks for these corrections._

I like the examples.

Referee: 2
Comments to the Author
I am pleased to recommend this manuscript for publication in Proceedings of the Royal Society Interface. The study of early warning indicators for biological transitions is an extremely active area of research. This paper reports the first study (to my knowledge) of the general trade-off between reliability and sensitivity of various indicators. Its introduction of an approximate method-based approach is a major contribution. The authors may wish to consider the following minor issues:

1. Line 41: “powerful way” is gratuitous

_"powerful" has been deleted_

2. Lines 61-63: I think some transcritical bifurcations can be “blue-sky” transitions; the claim should be qualified

_We apoloigize that the jargon is may not be used consistently in the field.  Transcritical bifurcations swap the stability of two fixed points, the fixed points to not appear/dissappear into the "blue".  We have removed the term_

3. Lines 74-76: I think the question of whether a noise-induced transition can be indicated is open. Particularly, the argument of Ditlevsen is based on a local approximation that is not expected to hold just prior to the transition; the author's may wish to qualify this claim. Also, “noise-induced” transition is ambiguous, here referring to the stochastic fluctuation between attractors, but more commonly referring to critical values in the magnitude of the fluctuations, as in Horsthemke and Lefever (1984).

_Excellent point, this has been clarified.  In particular, we note that the patterns put forward as warning signals for saddle node bifurcations, such as increasing variance or increasing autocorrelation, should not be expected to hold if the transition between attractors is driven purely by chance deviation._

4. Line 97: It is not obvious to be that autocorrelation from interpolation to a grid necessarily introduces artificial correlations. Either a reference or numerical example should be provided or the idea should be stated in a more conjectural mood.

5. Line 248: “This introduces a nuisance parameter known as the statistical significance criterion” is overly prejudicial. Differences of statistical philosophy are fine, but I think this goes over the line.

_This has been rephrased (we meant only in the technical sense as a parameter you must choose arbitrarily rather than estimate from data, but the point is well taken)_

6. Line 301: Typo

_The missing noun (citation to Dakos et al 2008) has been inserted,  thanks for catching this._