### 2012-06-01 fallacy.tex
fallacy explained more fully

%\documentclass[authoryear,5p]{elsarticle}
\documentclass[authoryear,review,12pt]{elsarticle}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
\usepackage{amsmath,amsfonts}
\usepackage{lineno}
\linenumbers
\usepackage{float}
\usepackage{subfigure}
\usepackage[pdftex]{color}
\definecolor{darkblue}{rgb}{0,0,0.5}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\usepackage[pdftex, colorlinks]{hyperref}
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\newcommand{\ud}{\mathrm{d}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\C}{\mathrm{Cov}}
\newcommand{\V}{\mathrm{Var}}
\newcommand{\cb}[1]{{\it \color{darkgreen} (#1)}}

%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{
\let\@oddhead\@empty
\let\@evenhead\@empty
\def\@oddfoot{\it \hfill\today}
\let\@evenfoot\@oddfoot}
\makeatother

\journal{\tiny } 
\begin{document}
\begin{frontmatter}
  \title{Early Warning Signals and the Prosecutor's Fallacy}
  \author[cpb]{Carl Boettiger\corref{cor1}}
  \ead{cboettig@ucdavis.edu}
  \author[esp]{Alan Hastings}
  %\author[info]{}
  %\author[davis]{}
  \cortext[cor1]{Corresponding author.}
  \address[cpb]{Center for Population Biology, 1 Shields Avenue, University of California, Davis, CA, 95616 United States.}
  \address[esp]{Department of Environmental Science and Policy, University of California, Davis}

 % \address[info]{ \\ 
  %              }

  \begin{abstract}

  Early warning signals have been proposed to forecast the possibility
  of a critical transition, such as the eutrophication of a lake,
  the collapse of a coral reef, or the end
  of a glacial period.  Because such transitions often occur on scales that
  can be difficult to approach by experimental manipulation, research
  has often relied on historical observations as a source of natural experiments.
  Here we examine a critical difference between selecting systems for
  study based on the fact that we have observed a critical transition
  and those systems for which we wish to forecast the approach of a
  transition. This difference arises by conditionally selecting systems
  known to experience a transition of some sort and failing to account
  for this bias this introduces -- a statistical error often known as the
  Prosecutor's Fallacy.  By analyzing simulated systems that have
  experienced transitions purely by chance, we reveal an elevated rate
  of false postives in common warning signal statistics. We further
  demonstrate a model-based approach which is less subject to this bias
  than these more commonly used summary statistics.
  In using historical examples of collapse to test and validate methods
  for early warning signals, care must be taken to avoid this fallacy,
  either through the use methods less sensitive to this effect or ideally
  by greater reliance on Experimental studies with replicates avoid this 
  pitfall entirely.  More such studies are needed to provide data suitable
  for evaluating methods for early warning.

  \end{abstract}

  \begin{keyword}
early warning signals \sep tipping point \sep alternative stable states \sep likelihood methods 
   \end{keyword}
 \end{frontmatter}

\section{Introduction}

\begin{quotation}
\noindent \emph{Mathematics \dots while assisting the trier of fact in the search of truth, must not cast a spell over him.}
-- California Supreme court, 1968.
\end{quotation}

\noindent In the case of \emph{People v. Collins} 1968, California Supreme
Court considered the evidence of an expert witness described by the
court as ``an instructor of mathematics at a state college'', which
concluded that the probability that a randomly selected individual
would match the description given by the victim would be less than 1 in
12 million~\citep{PeopleCollins1968}.  The prosecution had produced an
individual matching the prosecutor's detailed description, and convinced
by the mathematics of the instructor, the lower courts had found him
guilty.

The prosecution has only observed that the probability of seeing the
evidence ($E$) they produced given a random innocent individual ($I$),
$P(E|I)$ is very small.  This is not a sufficient basis from which
to conclude that the individual is indeed guilty, that is, that the
probability the individual is innocent given the evidence $P(I|E)$
is also very small. In a city with millions of people, there might
be several individuals who match the description of the evidence.
\footnote{Mathematically $P(E|I)$ need not equal $P(I|E)$, instead,
these expressions are related by Bayes theorem,

\begin{equation}
  P(E|I) = P(I|E) \frac{P(E)}{P(I)}
\end{equation} }

With little other evidence supporting the conviction, The California
Supreme Court reversed the decision, and the case became a widely
recognized example of the Prosecutor's Fallacy~\citep{Thompson1987}.
While DNA evidence has brought a rise in the number of probability-based
convictions in the courtroom, and a corresponding rise in the frequency
with which suspects are falsely accused, trial juries are not the
only ones who are susceptible to misinterpreting probabilistic evidence.
In this paper we explore how a similar misconception can arise from the
use of historical data to evaluate methods for detecting early warning
signals of critical transitions.

%\subsection*{Catastrophic Transitions}

Catastrophic transitions or tipping points, where a complex system
shifts suddenly from one state to another, have been implicated in
a wide array of ecological and global climate systems such as lake
ecosystems~\citep{Carpenter2011}, coral reefs~\citep{Mumby2007},
savannah~\citep{Kefi2007}, fisheries~\citep{Berkes2006}, and tropical
forests~\citep{Hirota2011}.  Recent research has begun to identify
statistical patterns commonly associated with these sudden catastrophic
transitions which could be used as an \emph{early warning sign} to
identify an approaching tipping point, which might provide managers time
to react to and avert an undesirable state shift~\citep{Scheffer2009, Lenton2011}.
An array of statistical patterns associated with tipping
point phenomena has been suggested for the detection of early warning
signals associated with such sudden transitions.  Two of the most commonly
used are a pattern of increasing variance~\citep{Carpenter2006} and a
pattern of increasing autocorrelation~\citep{vanNes2007}, which have
been tested in both experimental manipulation~\citep{Drake2010,
Carpenter2011, Veraart2011} and historical
observations~\citep{Livina2007,Dakos2008,Lenton2012,Ditlevsen2010,Guttal2008,Thompson2010}.

\subsection*{Testing patterns on historical data}

Historical examples of sudden transitions taken from the paleoclimate
record provide an important way to test and evaluate potential
leading indicator methods, and have been widely used for this purpose
\citep{Livina2007,Dakos2008,Lenton2012,Ditlevsen2010,Guttal2008,Thompson2010}.
Similarly, it has been suggested that data gathered from ecological
systems such as lakes that were monitored before they experienced sudden
eutriphication, or grasslands subjected to overgrazing, could contain
data that could help reveal when similar systems are approaching a
tipping point~\citep{Carpenter2011}.  This is a silver lining to the
clouds of past catastrophes; we may be able to learn from such examples
how to detect and avert sudden shifts in the future.

However, testing methods for early warning signals against historical
examples of transitions is suspetible to statistical mistakes that arise
from selecting data conditional on that data having already exhibited
a sudden transition.  A central tenent of early warning theory is that
the system in question is slowly approaching a tipping point that lies
some unknown distance away.  If nothing is done to remedy the situation,
this slow change will inevitably carry the system beyond the tipping
point, which introduces a sudden, rapid transition into an undesirable
state~\citep{Scheffer2009}. This process can be described mathematically
as a \emph{bifurcation}, in which a slowly changing parameter reaches
a critical value that causes the system stability to change.

Not all sudden transitions are caused by some guilty process slowly
driving the system over a tipping point, that is, the kind of process
that early warning signals are designed to detect.  Some systems may
experience such transitions purely by chance, leaving a stable state on
an extremely unlikely excursion that happens to stray to far from the
stable attractor~\citep[\emph{e.g.} ][consider this possibility in 
transitions that arise from analyzing historical climate record]{Ditlevsen2010, Lenton2011}. 
Like the evidence presented before the California Supreme
Court in 1968, the chance of observing such an ``innocent'' transition 
a priori may be very small, but when selected from a historical record
of many possible transitions, this possibility can no longer be ignored.

In principle, as such chance transitions do not involve the shrinking
eigenvalue that drives the critical slowing down and increasing variance
phenomena upon which several early warning signal statistics are based,
such data should not appear to display a warning signal. If this is the case,
then we can use historical data with more confidence when seeking to 
test early warning patterns, knowing that we will not mistake such examples
as cases of the bifurcation process we seek to detect.

Figure 1 shows a schematic illustrating critical transitions under
each of these scenarios.  In the left panel, the system experiences 
a bifurcation that forces it over the edge -- just the kind of data 
we would want to test an early warning signal.  In the right panel,
a similar-looking trajectory emerges from a simulation of a stable 
system.  While the simulation of the bifurcation scenario shown on
the left produces a similar transition every time, the transition 
shown on the right is somewhat less likely, occurring in only 1\% of
simulations.

\begin{figure}
  \begin{center}
    \subfigure[\textbf{True criminal} A bifurcation
    drives a transition into the undesirable state]{
    \includegraphics[width=.4\textwidth]{bifurcation.pdf}
    \label{fig:bifurcation} } \subfigure[\textbf{Innocent suspect} A
    transition to the alternative stable state has occurred purely by
    chance.]{ \includegraphics[width=.4\textwidth]{fallacytransition.pdf}
    \label{fig:fallacy} }
  \end{center} \caption{\textbf{The Prosecutor's Fallacy}. Plots of system
  state vs potential in two cases.  Potential is the negative integral
  of the mean of dynamical equation.  The potential function gives an
  intuitive picture of the stability of a system by imagining the curve as
  a surface on which a ball is free to bounce across -- wells correspond
  to stable points and peaks to unstable points.  Early warning signals
  are aimed at detecting systems which are slowly moving towards a tipping
  point or bifurcation, illustrated in the successively darker potential
  curves in panel~\subref{fig:bifurcation}.  An example trajectory from
  a simulation under this process shows the state of the system as the
  potential moves towards the bifurcation point. The state of the system
  in the simulation is given by the horizontal axis, while the vertical
  dimension is used to show time on the right-hand axis.  Some stable
  systems, panel~\subref{fig:fallacy}, may experience chance transitions
  to an alternate stable state without the presence of this slow change
  driving the system over the edge.  An example of the trajectory of
  such a chance transition is plotted over the potential.  While such
  events may be unlikely over short timescales, selecting historical
  examples of systems that have experienced sudden transitions increases
  the probability of accidentally using chance events depicted in
  panel~\subref{fig:fallacy} to test methods designed to detect the kind
  of process depicted in panel~\subref{fig:bifurcation}.} \label{fig:1}
\end{figure}

% Consider an equation

\section{Methods and Results}
We approach the problem by simulating a system that will not undergo the
saddle node bifurcation that underlies the usual early warning signal
indicators.  This simulation approach allows us to determine whether
examining historical events is a good way to test the utility of these
indicators.  We simulated 20,000 replicates of a stochastic individual-based
birth-death process with an Allee threshold~\citep{Courchamp2008}, which
arises from positive fitness effects at low densities.  Above the Allee
threshold the population returns to a positive equilibrium size, while
below the threshold the population decreases to zero in expectation. The
simulation uses the Gillespie algorithm to provide an exact implementation
of the master equation of the birth-death process,

\begin{align}
  \frac{dP(n,t)}{dt} &= b_{n-1} P(n-1,t) + d_{n+1}P(n+1,t) - (b_n+d_n) P(n,t)  \label{master} \\
    b_n &= \frac{e K n^2}{n^2 + h^2} \\
    d_n &= e n + a
\end{align}
a model with a linear death rate and density dependent birth rate
that drives the Allee effect at low densities and limits growth
at high densities.  The key feature of this model is the alternate
stable states introduced by this effect; other functional forms for
Eq.~\eqref{master} could serve equally well for these simulations~\citep[see
\emph{e.g.}][]{Scheffer2001}.  Though this system can be forced
through a saddle-node bifurcation by increasing the death rate, in
these simulations all parameters are held constant and no bifurcation
occurs.  Consequently we do not anticipate an early warning signal of
an approaching bifurcation.  The simulation starts from the positive
equilibrium population size.  Though the chance of a transition across
the Allee threshold in any given time step are small, given enough
time this system will eventually experience such a rare event driving
the population extinct.  We evolved each replicate 50,000 time steps,
sampling the system every 50 time steps.  In this time window 266 of the
1,000 replicates experience population collapse.  To keep the examples of
comparable sample size, we focus on a section of the data 500 time points
prior to the system approaching the transition.

To test whether selecting systems that have experienced spontaneous transitions
could bias the analysis towards false positive detection of early warning signals,
(the Prosecutor's Fallacy) we selected replicates conditional on having collapsed
in the simulations.
We then selected a window around each system that ended just before the
collapse, while the population values were still above the Allee threshold.
For each replicate, we calculated the most common early warning indicators, variance
and autocorrelation~\citep[\emph{e.g.}][]{Carpenter2006,Dakos2008,Scheffer2009}, 
around a moving window equal to half the length of that time series.

To test for the presence of a warning signal in these indicators we
computed values of Kendall's $\tau$ for both indicators for each of
the 266 replicates.  Kendall's $\tau$ takes values in $(-1, 1)$, and is
frequently used to identify an increasing trend ($\tau > 0 )$ in early
warning signals.~\citep{Dakos2008, Dakos2011}.  The distribution
of $\tau$ values observed across these replicates is shown in
Figure~\ref{fig:indicator}.  More positive values indicate stronger
trends. The common statistical significance test used for Kendall's $\tau$
assumes independence of the variables to achieve a normal distribution,
this isn't satisfied for indicators coming from a time series and
computed with a sliding window.  For a more robust comparison, we
compare the distribution of this statistic from all the simulations to
the distribution conditioned on experiencing a chance transition to the
alternative stable state, that is, from the replicates which are known
to crash.  To avoid an effect of sample size the time series are all chosen
to be the same length.  As these examples have been selected from a system
that is not approaching a bifurcation, an indicator that does not suffer
from the bias of Prosecutor's fallacy should have the same distribution
in both cases.   Annotated, parallelized code to replicate the analysis
and produce the figures is provided in the appendix.  (Currently at
\href{https://github.com/cboettig/earlywarning/blob/prosecutor/inst/examples/fallacy.md}{https://github.com/cboettig/earlywarning/blob/prosecutor/inst/examples/fallacy.md}).

\begin{figure}
  \begin{center}
    \includegraphics[width=6in]{indicators.pdf}
  \end{center}
  \caption{The distribution of the correlation statistic $\tau$ for two
  early warning indicators (variance, autocorrelation) on replicates
  conditionally selected for having collapsed by chance in simulations
  is shown in gray bars.  Solid lines indicate the estimated density of
  the statistic from a random sample of the simulations (not conditional
  on observing a transition). Positive values of $\tau$ correspond to
  a pattern of an indicator increasing with time; typically taken as
  evidence that a system is approaching a critical transition.  In these
  simulations, the pattern arises instead from the Prosecutor's fallacy
  of conditional selection.}
  \label{fig:indicator}
\end{figure}

For each of these replicates we also take a model-based approach, estimating 
parameters for an approximate linear model of the system approaching a
saddle node bifurcation.

\begin{equation}
  dX = \sqrt{ r_t } (\phi(r_t) - X_t) dt +
  \sigma\sqrt{\phi(r_t) } d B_t \label{LSN}
\end{equation}
where \( \phi(r_t) = \sqrt{r_t} +\theta \) is the equilibrium position.  
This model emerges from the canonical form of the saddle-node bifurcation,

\begin{equation}
  \frac{dX}{dt} = r - x^2
\end{equation}
where we introduce an explicit parameter $\theta$ to account for rescaling
to the units of the data,

\begin{equation}
   \frac{dX}{dt} = r - (x-\theta)^2
\end{equation}
and then consider the corresponding stochastic process where the noise is
proportional to the square root of the mean, as is the case for demographic
noise processes,

\begin{equation}
  \frac{dX}{dt} = r - (x-\theta)^2 + \sqrt{ \beta (r + (x-\theta)^2 } dB_t
\end{equation}
$r$ remains the bifurcation parameter which we imagine changing with time,
$r_t$, then linearizing the above model in neighborhood of
\(\hat x = \sqrt{r_t} +\theta =: \phi(r_t)\), we recover Equation~\eqref{LSN}.
These models can be both be fit to the data by maximum likelihood estimation
based on the equation for the evolution of their moments, as explained
in~\citet{Boettiger2012b}.  Though this approximation is more general than 
the original Markov process (as other models containing a saddle-node bifurcation
share the same approximation), it is also easier to estimate reliably from 
the data since it describes only the process near the equilibrium where the
data is available.  We assume that $r_t$ is varying at constant rate, 
according to the model

\begin{equation}
 r_t = m t + r_0 
\end{equation}
and estimate $m$, $r_0$, $\theta$, and $\beta$ by maximum likelihood on each
of the simulated datasets.  Estimates of $m < 0 $ are expected in systems 
approaching a bifurcation, while for stable systems $m$ should be approximately zero.

Of the 266 simulations, only six had estimates with machine precision different
from zero, the greatest magnitude among these was $-7.1 \times 10^{-12}$,
which would still correspond to a negligible change in $r_0$ over the 50,000 time steps
of the simulation.  Consequently, the model-based estimation shows no 
evidence of bias on data that has been selected conditional on collapse.

\section{Discussion}

Historical examples of systems that have undergone sudden transitions seem
to offer an invaluable opportunity to test methods for the detection of
early warning signals that could have foreseen the collapse.  However,
such data must be used with care, as the process of selecting systems
already known to have collapsed opens up the possibility that these
transitions were the result of rare chance events rather than examples
of the more deterministic tipping point behaviour we wish to detect.
When we think about forecasting potential collapse in a single system,
these purely-chance transitions may be rare enough to ignore.  When we
comb the historical record for any available data on sudden transitions,
our odds of finding these chance transitions may be much higher.

We have shown that systems which undergo such rare transitions look
statistically different from their counterparts that do not, even though
they are driven by the same stochastic process.  In particular, such
conditionally selected examples are more likely to show signs associated
with an early warning of an approaching tipping point, such as increasing
variance or increasing autocorrelation, as measured by Kendall's $\tau$.
This increases the risk of false positives -- cases in which a warning
signal being tested appears to have successfully detected an oncoming
tipping point, when in fact the example is falsely accused, coming
instead from a stable system. Figure 2 shows the extent of this bias,
where many of the chance crashes show values of $\tau$ that are 
significantly larger than those observed in the otherwise identical
replicates that did not experience a chance transition.

\subsection{Chance transitions are false positives for early warning signals}

It seems tempting to argue that the bias towards positive detection
in historical examples is valuable -- each of these systems did
indeed collapse, so the fact that they show increased probability
to exhibit warning signals could be taken as a successful detection.
Unfortunately this is not the case. At the moment the forecast is made,
these systems are no more likely to crash than their counterparts from
the sample that did not collapse, just as a fair coin is still a fair
coin after a string of consecutive ``heads.''  Taking a closer look at
the patterns involved we can begin to understand why common indicators
such as autocorrelation and variance can be misleading.

The farther the system gets from its stable point (while still remaining
in the local region), the more likely it is to draw a random step that
returns it towards the stable point, like a spring that is stretched
further and further. Despite this, there is always some probability that
it will move further still, and with enough possible examples to choose
from, we can be sure to see this happen occasionally.  Systems that manage
to extend this random walk far enough to cross the tipping point must do
so rather quickly -- with a string of events moving closer and closer to
the edge.  This pattern, clearly visible before the crashes in each of
the examples in Figure 1, produces a string of observations that appear
more highly autocorrelated (if we are sampling the system frequently
enough to catch the excursion at all) than we observe in the rest of the
fluctuations around the equilibrium.  Yet this autocorrelation comes from
a chance trajectory moving quickly \emph{away} from the stable state,
not from the critical slowing down pattern in the return times to the
stable state which precede a saddle-node bifurcation and motivate the
early warning signal.

This longer than expected excursion results in a higher than expected
variance in that window as well. Both variance and autocorrelation are
calculated using a moving window over the time-series, which allows
the method to pick out a pattern of change as the window moves along
the sequence. If this chance excursion that precedes the crash happens
to fill a significant part of the moving window, the resulting pattern
will tend to show an increase in autocorrelation or variance.  If the
chance excursion is relatively rapid compared to the frequency at which
the system is observed (spacing of the data) or the width of the moving
window, the excursion may not significantly alter the general pattern.
In this way, some of the events in which a crash is observed will
appear to present these statistical patterns of increased variance
or autocorrelation without being harbingers of approaching critical
transitions.

\subsection{Comparing to the model-based method}

In our experiment, the model-based estimate of early warning signals
appears more robust than the summary statistics, producing the same
estimates on both the conditionally selected replicates as on a random
sample of the replicates.  This is a consequence of the more rigid
specifications that come with a model-based approach -- the pattern
expected is less general than any increase in variance or autocorrelation,
but instead must be one that matches its approximation of the saddle-node
bifurcation. This observation highlights the difference between the
pattern driving the false positive trends in increasing variance and
increasing autocorrelation and the pattern anticipated in the saddle-node
model. This should not however be taken as evidence that the model-based
approach is immune to the bias of the Prosecutor's Fallacy.

The problem we highlight ultimately stems from the difficulty of having
only a single realization with which to examine a complex problem.
The only way to deal with this problem is through replication, as
can be done in an experimental system in laboratory manipulations
such as~\citet{Drake2010, Veraart2011} and at the scale of whole lake
ecosystems in~\citet{Carpenter2011}.  Historical data will not suffice,
more such replicated experiments are needed to test these patterns.
Experimental procedures avoid the hazard of the Prosecutor's fallacy by
generating a complete sample of replicates, rather than selecting a subset
of cases some larger historical sample where the risk that if selecting a
chance transition cannot be ignored.  Further, we note that the influence
of the Prosecutors Fallacy need not be restricted to the context of early
warning signals, but may arise any time we search historical data for
examples containing some feature of interest and ignore the statistical
effect that selection process introduces in our data.

 \section{Acknowledgements}
This research was supported by funding from NSF Grant EF 0742674 to AH
and a Computational Sciences Graduate Fellowship from the Department of
Energy grant DE-FG02-97ER25308 to CB. The authors thank M. Baskett, 
T.A. Perkins and N. Ross for helpful comments on earlier drafts of the
manuscript.

 \section{References}%bibliography
 \bibliography{fallacy}
 \end{document}