### 2013-07-11 inst/doc/sig-tests.Rmd
significance tests (use correct one-sided test)

```{r}
opts_chunk$set(comment=NA)
download.file("https://raw.github.com/cboettig/earlywarning/resubmission/inst/doc/Figure1.csv", dest="Figure1.csv", method="curl")
dat <- read.csv("Figure1.csv")
```

There should be some clever way in plyr or data.table to do these combinations automatically, but anyway:

Focusing on the OU model, we see that autocorrelation is not sufficiently different:

```{r}
x <- subset(dat, variable=="Autocorrelation" & data=="null" & model == "OU")$value
y <- subset(dat, variable=="Autocorrelation" & data=="conditional" & model == "OU")$value
wilcox.test(x,y, alternative="less") # x - y < 0 since y bigger than x
ks.test(x,y, alternative="greater") # CDF x > then CDF y since more weight to the left
```

Variances are more different:

```{r}
x <- subset(dat, variable=="Variance" & data=="null" & model == "OU")$value
y <- subset(dat, variable=="Variance" & data=="conditional" & model == "OU")$value
wilcox.test(x,y, alternative="less") # x - y < 0 since y bigger than x
ks.test(x,y, alternative="greater") # CDF x > then CDF y since more weight to the left
```

Clearer for Allee case:

```{r}
x <- subset(dat, variable=="Variance" & data=="null" & model == "Allee")$value
y <- subset(dat, variable=="Variance" & data=="conditional" & model == "Allee")$value
wilcox.test(x,y, alternative="less") # x - y < 0 since y bigger than x
ks.test(x,y, alternative="greater") # CDF x > then CDF y since more weight to the left
```

We opt not to include a comparison statistic since it is unclear to me what is most appropriate.  Wilcox test is a bit of a strange choice given the skew of the conditional distribution.  Kolmogorov-Smirnov is slightly more natural, but really hard to defend either one in this context, as distributions of Kendall's $\tau$ statistics is already a somewhat curious distribution - ideally we should consider how it should actually be distributed given the dynamical process as the null (which is what we generate numerically).