### 2014-11-20 prosecutor-fallacy-code/warning_signal_slides.Rmd

```{r }
#opts_knit$set(upload.fun = socialR::flickr.url)

require(ggplot2)
require(reshape2)
theme_publish <- theme_set(theme_bw(14))

lower <- 0
upper <- 700
x <- seq(lower, upper, length.out=100)
e <- 0.5 
K <- 1000
h <- 200

a <- seq(120, 220, by=20)

d0 <- sapply(a, function(a){
b <- function(n)  (0.5 * K * n ^ 2) / (n ^ 2 + h ^ 2)
d <- function(n) e * n + a
f <- function(n) b(n) - d(n)
bx <- sapply(x, b) 
dx <- sapply(x, d) 
growth <- bx-dx
})

d1 <- melt(data.frame(State=x, d0), id="State")
ggplot(d1) + 
  geom_line(aes(State, value, lty=variable), lwd=1) + 
  ylab("growth rate")

```

```{r }
a <- 180
u1 <- -1 * sapply(x, function(x) integrate(f, lower, x)$value)
a <- 190
u2 <- -1 * sapply(x, function(x) integrate(f, lower, x)$value)
a <- 200
u3 <- -1 * sapply(x, function(x) integrate(f, lower, x)$value)
```

```{r chance_crash_data} 
load("data/zoom.rda")
dat <- subset(zoom, reps==2)
#dat <- subset(dat, value > 250)
a <- dat$time - min(dat$time)
a <- a/max(a)*(34000-19000) + 19000
d2 <- data.frame(x = dat$value, value = a)

```

```{r Figure1b, tidy=FALSE}
df <- data.frame(x=x, value=u1)
df$variable = "stable potential"
d2$variable = "stable potential"
df$panel <- "Potential Energy"
d2$panel <- "Time of Sample Trajectory"
d <- rbind(df, d2)
p1 <- ggplot(data = d, mapping = aes(x = x, y = value)) + 
  facet_grid(panel ~ ., scale = "free") +
  layer(data = d2, geom = "point", size=1) +
  layer(data = df,  mapping = aes(lty=variable),  geom = c("line")) +
  geom_vline(aes(xintercept=250), lty=4) +
  coord_cartesian(ylim=c(18000, 35000)) + 
  xlab("State") + ylab("arbitrary units") +
  opts(title = "(b) Trajectory of a chance \n transition from a stable system")

write.csv(d, "Figure1_panelb.csv")
```

simulate an example with a bifurcation for comparison.

```{r crash_sim, cache=TRUE}
  library(populationdynamics)
	T<- 5000
	n_pts <- 50000
	pars = c(Xo = 500, e = 0.5, a = 180, K = 1000, h = 200,
    i = 0, Da = .5, Dt = 0, p = 2)
	sn <- saddle_node_ibm(pars, times=seq(0,T, length=n_pts), reps=1)
  X <- sn$x1
  X <- X[X > 0]
  times <- 1:length(X)*T/n_pts
  a <- times - min(times) 
  a <- a/max(a)*(34000-19000) + 19000
  d3 <- data.frame(x=X, value=a)
````

```{r Figure1c, tidy=FALSE}
df <- data.frame(x=x, "stable potential"=u1, "deteriorating"=u2, "critical"=u3)
d1 <- melt(df, id="x")
d1$panel <- "Potential Energy"
d3$panel <- "Time of Sample Trajectory"
d3$variable <- d3$x # match col numbers to rbind
d <- rbind(d1, d3)
p2 <- ggplot(data = d, mapping = aes(x = x, y = value)) + 
  facet_grid(panel ~ ., scale = "free") +
  layer(data = d3, geom = "point", size=1) +
  layer(data = d1,  mapping = aes(lty=variable), geom = c("line")) +
  geom_vline(aes(xintercept=250), lty=4) +
  coord_cartesian(ylim=c(18000, 35000)) + xlab("State") + ylab("arbitrary units") +
  opts(title="(c) Trajectory during an \n underlying parameter shift")
write.csv(d, "Figure1_panelc.csv")
```

```{r Figure1_upload, fig.width=15}
require(grid)
pushViewport(viewport(layout = grid.layout(1, 3)))
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
print(p0, vp = vplayout(1, 1))
print(p1, vp = vplayout(1, 2))
print(p2, vp = vplayout(1, 3))
```

```{r Figure1_orig, fig.width=15, dev="Cairo_pdf"}
require(grid)
pushViewport(viewport(layout = grid.layout(1, 3)))
vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
print(p0, vp = vplayout(1, 1))
print(p1, vp = vplayout(1, 2))
print(p2, vp = vplayout(1, 3))
```